---
title: "Foundations of probability in R"
output: html_document
---


#Chapter 1 The binomial distribution 

```{r}
# Generate 10 separate random flips with probability .3
rbinom(10,1,.3)
```

```{r}
# Generate 100 occurrences of flipping 10 coins, each with 30% probability
rbinom(100, 10, .3)
```

dbinom --> that it is exactly that amount of times i get head

pbinom --> the change of getting up to a given amount of heads


```{r}
# Calculate the probability that 2 are heads using dbinom
dbinom(2, 10, .3)


# Confirm your answer with a simulation using rbinom
mean(rbinom(10000, 10, .3)==2)
```

```{r}
# Calculate the probability that at least five coins are heads
1 - pbinom(4, 10, .3)

# Confirm your answer with a simulation of 10,000 trials
mean(rbinom(10000, 10, .3) >= 5)
```

```{r}
# Here is how you computed the answer in the last problem
mean(rbinom(10000, 10, .3) >= 5)

# Try now with 100, 1000, 10,000, and 100,000 trials
mean(rbinom(100, 10, .3) >= 5)

mean(rbinom(1000, 10, .3) >= 5)
mean(rbinom(10000, 10, .3) >= 5)
mean(rbinom(100000, 10, .3) >= 5)
```

```{r}
# Calculate the expected value using the exact formula
25*0.3

# Confirm with a simulation using rbinom
mean(rbinom(10000, 25, .3))
```


```{r}
# Calculate the variance using the exact formula
25*0.3*0.7

# Confirm with a simulation using rbinom
var(rbinom(10000, 25, .3))
```


#Chapter 2 Laws of probability

Probability that A and B happens 

```{r}
# Simulate 100,000 flips of a coin with a 40% chance of heads
A <- rbinom(100000, 1, 0.4)

# Simulate 100,000 flips of a coin with a 20% chance of heads
B <- rbinom(100000, 1, 0.2)

# Estimate the probability both A and B are heads
mean(A & B) 
```


```{r}
# You've already simulated 100,000 flips of coins A and B
A <- rbinom(100000, 1, .4)
B <- rbinom(100000, 1, .2)

# Simulate 100,000 flips of coin C (70% chance of heads)
C= rbinom(100000, 1, 0.7)

# Estimate the probability A, B, and C are all heads

mean(A & B & C)
```



probability of A or B happens 

```{r}
# Simulate 100,000 flips of a coin with a 60% chance of heads
A <- rbinom( 100000, 1, 0.6)

# Simulate 100,000 flips of a coin with a 10% chance of heads
B <- rbinom( 100000, 1, 0.1)

# Estimate the probability either A or B is heads

mean(A|B)
```


```{r}
# Use rbinom to simulate 100,000 draws from each of X and Y
X <- rbinom(100000, 10, .6)
Y <- rbinom(100000, 10, .7)

# Estimate the probability either X or Y is <= to 4
mean(X <= 4 | Y <= 4)

# Use pbinom to calculate the probabilities separately
prob_X_less <- pbinom(4, 10, .6)
prob_Y_less <- pbinom(4, 10, .7)

# Combine these to calculate the exact probability either <= 4
prob_X_less + prob_Y_less - prob_X_less * prob_Y_less
```

- Multiplying random variables 

if we multiply by af constant K

```{r}
# Simulate 100,000 draws of a binomial with size 20 and p = .1
X <- rbinom( 100000, 20, .1)

# Estimate the expected value of X
mean(X)

# Estimate the expected value of 5 * X

mean(5*X)
```


```{r}
# X is simulated from 100,000 draws of a binomial with size 20 and p = .1
X <- rbinom(100000, 20, .1)

# Estimate the variance of X
var(X)

# Estimate the variance of 5 * X
var(5*X)
```


- Adding two random variables 

If the variables are independent --> E[X+Y]= E[X]+ E[Y] (Also if X and Y are not independent )

Same the Var[X+Y]= Var[X]+ Var[Y]

```{r}
# Simulate 100,000 draws of X (size 20, p = .3) and Y (size 40, p = .1)
X <- rbinom(100000, 20, .3)
Y <- rbinom(100000, 40, .1)

# Estimate the expected value of X + Y

mean(X+Y)
```


```{r}
# Simulation from last exercise of 100,000 draws from X and Y
X <- rbinom(100000, 20, .3) 
Y <- rbinom(100000, 40, .1)

# Find the variance of X + Y

var(X+Y)

# Find the variance of 3 * X + Y
var(3*X+Y)
```


#Chapter 3 Bayesian statistics

```{r}
# Simulate 50000 cases of flipping 20 coins from fair and from biased
fair <- rbinom(50000, 20, .5)
biased <- rbinom(50000, 20, .75)

# How many fair cases, and how many biased, led to exactly 11 heads?
fair_11 <- sum(fair==11)
biased_11 <- sum(biased == 11)

# Find the fraction of fair coins that are 11 out of all coins that were 11
fair_11/(fair_11+biased_11)
```


```{r}
# Simulate 50000 cases of flipping 20 coins from fair and from biased
fair <- rbinom(50000, 20, .5)
biased <- rbinom(50000, 20, .75)

# How many fair cases, and how many biased, led to exactly 16 heads?
fair_16 <- sum(fair==16)
biased_16 <- sum(biased==16)

# Find the fraction of fair coins that are 16 out of all coins that were 16
fair_16/(fair_16+ biased_16)
```


- Now there is no longer a 50/50 from the beginning (80/20 now)

```{r}
# Simulate 8000 cases of flipping a fair coin, and 2000 of a biased coin
fair_flips <- rbinom(8000, 20, .5)
biased_flips <-rbinom(2000, 20, .75)

# Find the number of cases from each coin that resulted in 14/20
fair_14 <- sum(fair_flips==14)
biased_14 <- sum(biased_flips == 14)

# Use these to estimate the posterior probability

fair_14/(fair_14+ biased_14)
```


```{r}
# Simulate 80,000 draws from fair coin, 10,000 from each of high and low coins
flips_fair <- rbinom(80000, 20 ,.5)
flips_high <- rbinom(10000, 20, .75)
flips_low <- rbinom(10000, 20, .25)

# Compute the number of coins that resulted in 14 heads from each of these piles
fair_14 <- sum(flips_fair== 14)
high_14 <- sum(flips_high== 14)
low_14 <- sum(flips_low== 14)

# Compute the posterior probability that the coin was fair

fair_14/(fair_14+ high_14+ low_14)
```

Bayes' theoreme

```{r}
# Use dbinom to calculate the probability of 11/20 heads with fair or biased coin
probability_fair <- dbinom(11, 20, .5)
probability_biased <- dbinom(11, 20, .75)

# Calculate the posterior probability that the coin is fair
probability_fair / (probability_fair + probability_biased)
```


```{r}
# Find the probability that a coin resulting in 14/20 is fair
probability_fair <- dbinom(14, 20, .5)
probability_biased <- dbinom(14, 20, .75)

probability_fair / (probability_fair + probability_biased)


# Find the probability that a coin resulting in 18/20 is fair
probability_fair1 <- dbinom(18, 20, .5)
probability_biased1 <- dbinom(18, 20, .75)

probability_fair1 / (probability_fair1 + probability_biased1)
```

```{r}
# Use dbinom to find the probability of 16/20 from a fair or biased coin
probability_16_fair <- dbinom(16, 20, .5)
probability_16_biased <- dbinom(16, 20, .75)

# Use Bayes' theorem to find the posterior probability that the coin is fair
probability_16_fair*0.99/ (probability_16_fair*0.99+ probability_16_biased*0.01)
```


#Chapter 4 Normal distribution



```{r}
# Draw a random sample of 100,000 from the Binomial(1000, .2) distribution
binom_sample <- rbinom(100000, 1000, .2)

# Draw a random sample of 100,000 from the normal approximation
normal_sample <-  rnorm(100000, 200, sqrt(160))

# Compare the two distributions with the compare_histograms function
compare_histograms(binom_sample, normal_sample)
```


```{r}
# Simulations from the normal and binomial distributions
binom_sample <- rbinom(100000, 1000, .2)
normal_sample <- rnorm(100000, 200, sqrt(160))

# Use binom_sample to estimate the probability of <= 190 heads

mean(binom_sample<=190)

# Use normal_sample to estimate the probability of <= 190 heads

mean(normal_sample <= 190)

# Calculate the probability of <= 190 heads with pbinom

pbinom(190, 1000, .2)

# Calculate the probability of <= 190 heads with pnorm

pnorm(190, 200, sqrt(160))
```



```{r}
# Draw a random sample of 100,000 from the Binomial(10, .2) distribution
binom_sample <- rbinom(100000, 10, .2)

# Draw a random sample of 100,000 from the normal approximation
normal_sample <- rnorm(100000, 2, sqrt(1.6))

# Compare the two distributions with the compare_histograms function

compare_histograms(binom_sample, normal_sample)
```


Poisson distribution 

```{r}
# Draw a random sample of 100,000 from the Binomial(1000, .002) distribution
binom_sample <- rbinom(100000, 1000, .002)

# Draw a random sample of 100,000 from the Poisson approximation
poisson_sample <- rpois(100000, 2)

# Compare the two distributions with the compare_histograms function
compare_histograms(binom_sample, poisson_sample)
```


```{r}
# Simulate 100,000 draws from Poisson(2)
poisson_sample <- rpois(100000, 2)

# Find the percentage of simulated values that are 0
mean(poisson_sample==0)

# Use dpois to find the exact probability that a draw is 0
dpois(0, 2)
```


```{r}
# Simulate 100,000 draws from Poisson(1)
X= rpois(100000, 1)

# Simulate 100,000 draws from Poisson(2)
Y= rpois(100000, 2)

# Add X and Y together to create Z
Z=X+Y

# Use compare_histograms to compare Z to the Poisson(3)

compare_histograms(Z, rpois(100000, 3))
```


The geometric distribution

-use Which function to see which position the first head has. 
- expected vaslues (1/P)-1


```{r}
# Simulate 100 instances of flipping a 20% coin
flips <- rbinom(100, 1, .2)

# Use which to find the first case of 1 ("heads")
which(flips==1)[1]
```


```{r}
# Existing code for finding the first instance of heads
which(rbinom(100, 1, 0.2) == 1)[1]

# Replicate this 100,000 times using replicate()
replications <- replicate(100000, which(rbinom(100, 1, 0.2) == 1)[1])

# Histogram the replications with qplot
qplot(replications)
```

```{r}
# Replications from the last exercise
replications <- replicate(100000, which(rbinom(100, 1, .2) == 1)[1])

# Generate 100,000 draws from the corresponding geometric distribution
geom_sample <- rgeom(100000, .2)

# Compare the two distributions with compare_histograms
compare_histograms(replications, geom_sample)
```

```{r}
# Find the probability the machine breaks on 5th day or earlier
pgeom(4, .1)

# Find the probability the machine is still working on 20th day
1 - pgeom(19, .1)
```


```{r}
# Calculate the probability of machine working on day 1-30
still_working <- 1-pgeom(0:29, .1)

# Plot the probability for days 1 to 30
qplot(1:30, still_working)
```
